---
title: "final_sna"
author: "Phong Duong, Nienke Visscher, Frederick Sims"
date: "2024-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

# Libraries

```{r}
library(tidyverse)
library(igraph)
```

# Import Data

```{r}
data <- read_csv("migration_data.csv") %>% 
  rename(source = "Country Origin Name",
         target = "Country Dest Name",
         migration_2000 = "2000 [2000]") %>% 
  select(source, target, migration_2000) %>% 
  mutate(
    source = gsub(x = source, pattern = "\\.", replacement = ""),
    target = gsub(x = target, pattern = "\\.", replacement = ""),
    migration_2000 = as.numeric(migration_2000)) %>% 
  filter(migration_2000 > 0) %>% 
  group_by(source,target) %>% 
  summarise(migration_2000 = sum(migration_2000))
data %>% head()
```

```{r}
g <- graph_from_data_frame(data, directed = TRUE) %>% 
  simplify(.) 
E(g)$weight <- data$migration_2000
g
```

```{r}
par(mar=c(0,0,0,0))
plot(g, directed = TRUE, weighted = TRUE,
     edge.arrow.size = 0.2,
     edge.color = adjustcolor("grey30", 0.2),
     vertex.color = ifelse(degree(g) > 10, "orange", "grey50"),
     vertex.size = 4,
     vertex.label = "")
```

# Propagation Simulation

We want to simulate the propagation of information on this network using the SIR model assuming that an infected person can transmit the information with probability β to a susceptible person on the network.

## Epidemic Threshold

a)  Find the theoretical epidemic threshold β for your network for the information to reach a significant number of nodes.

```{r}
mu <- 0.1
betac <- mu*mean(degree(g))/(mean(degree(g)^2))
betac
```

## Simulation

b)  Assuming that randomly-selected 1% of the nodes of your network knows about the information, simulate the SIR model below and above that threshold and plot the number of infected people as a function of β.

```{r}
sim_sir <- function(g,beta,mu,seeds){
  state <- rep(0,vcount(g)) #initial state of the simulation
  state[seeds] <- 1 #infect the seeds
  t <- 0
  table <- data.frame(t=0,inf=seeds)
  while(sum(state==1)>0){
    t <- t + 1
    ## I -> R
    infected <- which(state==1) #get them
    
    # generate a random value for every infected, if it's < mu, let the node recover.
    state[infected] <- ifelse(runif(length(infected)) < mu,2,1)
    
    ## S -> I
    infected <- which(state==1)
    susceptible <- which(state==0) #get them
    
    ###Get contacts with weighted edges
    if(sum(state ==0) >0 & sum(state == 1) > 0) { #stop generating infection once all are recovered
    contacts <- adjacent_vertices(g, infected) %>% #find neighbors of infected vertices
      unlist() %>% 
      names() %>% #get the names of all the nodes
      sub(patter = "\\.", replacement =  "%") %>% 
      data.frame() %>% 
      separate(".", c("vi", "vj"), sep = "%") %>% #separate source and target nodes in pairs to id 
      mutate(vi_id = lapply(vi, function(x){id <- which(V(g)$name == x)}) %>% unlist(), #id source node
             vj_id = lapply(vj, function(x){id <- which(V(g)$name == x)}) %>% unlist(), #id target node
             weight = E(g)[vi_id %->% vj_id]$weight) %>% #add weight of the link
      filter(vj_id %in% susceptible) %>% #filter only susceptible nodes 
      #calculate beta_ij weighted by immigrant amount
      group_by(vj_id) %>% 
      mutate(sum_i_log_wij = sum(log(weight))) %>% 
      ungroup() %>%
      mutate(beta_ij = ifelse(sum_i_log_wij == 0, 0 ,beta * log(weight) / sum_i_log_wij))
    
    ###Generate infection
    new_infected <- lapply(contacts$beta_ij, function(x){inf <- runif(1) < x}) %>% unlist()
    new_infected <- contacts$vj_id[new_infected]
    new_infected <- new_infected[!duplicated(new_infected)]
    
    ### Append nodes newly infected into the table
    if(length(new_infected)>0){
      state[new_infected] <- 1
      table <- rbind(table,data.frame(t,inf=new_infected))
    }
    }
  }
  table
}
```

### Above Threshold

```{r}
set.seed(230)
# Sample 1% of the network as initial infectious seed
seeds <- sample(1:vcount(g),0.01*vcount(g))
# Generate a simlulation of the SIR model with beta above threshold
realization_at <- sim_sir(g,beta = 0.3,mu=0.1,seeds) 
realization_at
```

```{r}
realization_at <- realization_at %>% 
  group_by(t) %>% summarize(ninf=n()) 
realization_at %>% 
  ggplot(aes(x=t,y=ninf)) + geom_line()
```

### Below threshold

```{r}
set.seed(230)
# Sample 1% of the network as initial infectious seed
seeds <- sample(1:vcount(g),0.01*vcount(g))
# Generate a simlulation of the SIR model with beta below threshold
realization_bt <- sim_sir(g,beta = 0.0002, mu=0.01,seeds) 
realization_bt
```

```{r}
realization_bt <- realization_bt %>% 
  group_by(t) %>% summarize(ninf=n()) 
realization_bt %>% 
  ggplot(aes(x=t,y=ninf)) + geom_line()
```

### Different Betas

```{r}
set.seed(230)
results <- map_dfr(seq(0,0.2,0.005), # beta range
        \(beta){ seeds <- sample(1:vcount(g),vcount(g)*0.01)
        realization <- sim_sir(g,beta,mu,seeds) # make a realization for each beta
        data.frame(beta,ninf=nrow(realization)) # create dataframe row
        })
results %>% ggplot(aes(x=beta,y=ninf))+ geom_point()+
  geom_vline(xintercept = betac,linetype=2)
```

## Better 1% seeds

c)  Choose a β above β~c~. Using centrality, communities or any other metric, find a better set of 1% of seeds in the network so we get more infected people than the random case. Measure the difference of your choice with the random case as:

```{r}
set.seed(230)
beta_at <- 0.5
#target at random
seeds_random <- sample(1:vcount(g),vcount(g)*0.01)
realization_random <- sim_sir(g,beta = beta_at,mu=0.1,seeds_random) %>%
  group_by(t) %>% summarize(ninf=n())
#target by closeness
seeds_closeness <- closeness(g) %>% enframe(name = "inf",value="closeness") %>%
  slice_max(closeness,n= round(vcount(g)*0.01))
realization_closeness <- sim_sir(g,beta = beta_at,mu=0.1,which(V(g)$name == seeds_closeness$inf)) %>%
  group_by(t) %>% summarize(ninf=n())
#target by betweenness
seeds_between <- betweenness(g) %>% enframe(name = "inf",value="betweenness") %>%
  slice_max(betweenness,n= round(vcount(g)*0.01))
realization_between <- sim_sir(g,beta = beta_at,mu=0.1,which(V(g)$name == seeds_between$inf)) %>%
  group_by(t) %>% summarize(ninf=n())
#target using page rank
seeds_pagerank <- page_rank(g)$vector %>% enframe(name = "inf",value="page_rank") %>%
  slice_max(page_rank,n= round(vcount(g)*0.01))
realization_pagerank <- sim_sir(g,beta = beta_at,mu=0.1,which(V(g)$name == seeds_pagerank$inf)) %>%
  group_by(t) %>% summarize(ninf=n())

ggplot() + geom_line(data=realization_random,aes(x=t,y=ninf,col="Target at random"))+
  geom_line(data=realization_closeness,aes(x=t,y=ninf,col="Target by closeness")) +
  geom_line(data=realization_between,aes(x=t,y=ninf,col="Target by betweenness")) +
  geom_line(data=realization_pagerank,aes(x=t,y=ninf,col="Target by page rank"))
```

-   The difference in the total number of infected people

```{r}
realization_random$ninf %>% sum()
realization_between$ninf %>%  sum()
realization_closeness$ninf %>%  sum()
realization_pagerank$ninf %>% sum()
```

-   The difference in the time of the peak of infection (when most infections happen).

```{r}
max(realization_random$ninf)
max(realization_between$ninf)
max(realization_closeness$ninf)
max(realization_pagerank)
```

## Information stoppers

### 5% randomly

d)
Choose those 5% randomly in the network. Simulate the SIR model above using beta~c 1% of the remaining nodes as seeds. Choose those seeds randomly.

```{r}
beta_at <- 0.1
```


```{r}
set.seed(230)
vaccinated_random <- sample(1:vcount(g),vcount(g)*0.05)
gp_random <- delete_vertices(g,vaccinated_random)
seeds <- sample(1:vcount(gp_random),vcount(gp_random)*0.01)
vaccination_random <- sim_sir(gp_random,beta = beta_at,mu=0.1,seeds) %>%
  group_by(t) %>% summarize(ninf=n())
```

### Vaccination according to their centrality

Choose those 5% according to their centrality. Simulate the SIR model above using 1% of the remaining nodes as seeds. Choose those seeds randomly.

```{r}
# immunization by closeness
set.seed(4785)
vaccinated_closeness <- closeness(g) %>% enframe(name = "inf",value="closeness") %>%
  slice_max(closeness,n = round(vcount(g)*0.05))
gp_closeness <- delete_vertices(g,vaccinated_closeness$inf)
seeds <- sample(1:vcount(gp_closeness),vcount(gp_closeness)*0.01)
vaccination_targeted_closeness <- sim_sir(gp_closeness,beta = beta_at,mu=0.1, seeds) %>% 
  group_by(t) %>%
  summarize(ninf=n())

# immunization by betweenness
set.seed(4575)
vaccinated_between <- betweenness(g) %>% enframe(name = "inf",value="betweenness") %>%
  slice_max(betweenness,n=round(vcount(g)*0.05))
gp_between <- delete_vertices(g,vaccinated_between$inf)
seeds <- sample(1:vcount(gp_between),vcount(gp_between)*0.01)
vaccination_targeted_between <- sim_sir(gp_between,beta = beta_at,mu=0.1,seeds) %>%
  group_by(t) %>% summarize(ninf=n())

# immunization by page rank
set.seed(365)
vaccinated_pagerank <- page_rank(g)$vector %>% enframe(name = "inf",value="page_rank") %>%
  slice_max(page_rank,n=round(vcount(g)*0.05))
gp_pagerank <- delete_vertices(g,vaccinated_pagerank$inf)
seeds <- sample(1:vcount(gp_pagerank),vcount(gp_pagerank)*0.01)
vaccination_targeted_pagerank <- sim_sir(gp_pagerank,beta = beta_at,mu=0.1,seeds) %>%
  group_by(t) %>% summarize(ninf=n())
```

Measure the difference between both cases as you did in part c).

```{r}
ggplot() + 
  geom_line(data=vaccination_random,aes(x=t,y=ninf,col="Vacc. Random"))+
  geom_line(data=vaccination_targeted_closeness,aes(x=t,y=ninf,col="Vacc. Targeted Closeness"))+ 
  geom_line(data=vaccination_targeted_between,aes(x=t,y=ninf,col="Vacc. Random Between"))+
  geom_line(data=vaccination_targeted_pagerank,aes(x=t,y=ninf,col="Vacc. Targeted PageRank"))
```

-   The difference in the total number of infected people

```{r}
vaccination_random$ninf %>% sum()
vaccination_targeted_closeness$ninf %>%  sum()
vaccination_targeted_between$ninf %>% sum()
vaccination_targeted_pagerank$ninf %>%  sum()
```

-   The difference in the time of the peak of infection (when most infections happen).

```{r}
max(vaccination_random$ninf)
max(vaccination_targeted_closeness$ninf)
max(vaccination_targeted_between$ninf)
max(vaccination_targeted_pagerank$ninf)
```

## Effect of vaccination

Comment on the relationship between the findings in part c) and d) using the same type of centrality for the 1% in part c) and 5% in part d).